---
title: "Phlox Project - Geralin, Natascha, Jala"
output: 
  html_document: 
    keep_md: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Installing Libraries if Necessary (for non-frequent R users)
```{r}
if (!require("tidyverse")) install.packages('tidyverse')
if (!require("RColorBrewer")) install.packages('RColorBrewer')
if (!require("paletteer")) install.packages('paletteer')
if (!require("janitor")) install.packages('janitor')
if (!require("here")) install.packages('here')
if (!require("naniar")) install.packages('naniar')
if (!require("stringr")) install.packages('stringr')
if (!require("shiny")) install.packages('shiny')
if (!require("shinydashboard")) install.packages('shinydashboard')
```

#Library
```{r message=FALSE, warning=FALSE, comment=NA}
library(tidyverse)
library(RColorBrewer)
library(paletteer)
library(janitor)
library(here)
library(naniar)
library(stringr)
library(shiny)
library(shinydashboard)
```

#Loading Original Data Sets and Cleaning Names
The following data sets examine the pollinators, colors, and sizes associated with different species of flowers within the phlox family. Source: Landis, Jacob B. et al. (2018), Evolution of floral traits and impact of reproductive mode on diversification in the phlox family (Polemoniaceae), Molecular Phylogenetics and Evolution, Article-journal, https://doi.org/10.1016/j.ympev.2018.06.035
[Data:](https://datadryad.org/stash/dataset/doi:10.5061/dryad.2710pk5)
```{r message=FALSE, warning=FALSE, comment=NA}
flower_sizes_original <- read_csv(here("Plant+Pollinators Data", "Flower size measurements.csv")) %>%
  janitor::clean_names()
colors_and_pollinators_original <- read_csv(here("Plant+Pollinators Data", "Pollinator and flower color.csv")) %>%
  janitor::clean_names()
```

#Flower Sizes Data
With the first data set, we'll observe the lengths and widths of each species' corollas, removing any NAs recorded for these variables.
```{r message=FALSE, warning=FALSE, comment=NA}
flower_sizes <- flower_sizes_original %>% 
  select(herbarium_accession_sheet, species, flower_number, corolla_length_cm, corolla_width_throat_cm) %>% 
  filter(corolla_length_cm != "NA", corolla_width_throat_cm != "NA")
flower_sizes
```
```{r message=FALSE, warning=FALSE, comment=NA}
glimpse(flower_sizes)
```
```{r message=FALSE, warning=FALSE, comment=NA}
flower_sizes %>%
  miss_var_summary() #This is to confirm that we have removed all NAs represented as "NA" from the size columns, as well as to get a count of any other NAs in the remaining columns
```

#Calculate the means, standard deviations and length to width ratios of the corollas of each species:
```{r message=FALSE, warning=FALSE, comment=NA}
flower_sizes_calculated <- flower_sizes %>% 
  group_by(species) %>% 
  summarise(mean_corolla_length_cm = mean(corolla_length_cm, na.rm = TRUE),
         std_dev_corolla_length = sd(corolla_length_cm, na.rm = TRUE),
         mean_corolla_width_throat_cm = mean(corolla_width_throat_cm, na.rm = TRUE),
         std_dev_corolla_width_throat = sd(corolla_width_throat_cm, na.rm = TRUE),
         mean_length_width_ratio = mean_corolla_length_cm/mean_corolla_width_throat_cm)
flower_sizes_calculated #summary of sizes for each species, with means, standard deviations, and ratios calculated
```

```{r message=FALSE, warning=FALSE, comment=NA}
n_distinct(flower_sizes_calculated$species)
```

There are 396 species in our flower sizes data sheet. Object `flower_sizes_calculated` should be used for the merge. 

#Pollinator and Color Data
With our second data set, we'll observe the different colors and pollinators associated with each species. Note: this data is not yet clean.
```{r message=FALSE, warning=FALSE, comment=NA}
colors_and_pollinators <- colors_and_pollinators_original %>% 
  rename(color_source="source", pollinator_source="source_1") %>% 
  filter(pollinator != "NA")
colors_and_pollinators 
```

```{r message=FALSE, warning=FALSE, comment=NA}
glimpse(colors_and_pollinators)
```

```{r message=FALSE, warning=FALSE, comment=NA}
n_distinct(colors_and_pollinators$species)
```

```{r message=FALSE, warning=FALSE, comment=NA}
miss_var_summary(colors_and_pollinators)
```

There are 129 distinct species in the `colors_and_pollinators` data and 397 in the `flower_size` data. We will use inner_join so that we are only working with species that have data in both data sets.

#The Merge
```{r message=FALSE, warning=FALSE, comment=NA}
phlox_merge <- inner_join(flower_sizes_calculated, colors_and_pollinators, by = "species")
phlox_merge
```
#Final count of the number of species we are working with: 
```{r}
n_distinct(phlox_merge$species)
```

After joining the data sheets, we have 145 species that we are working with. 

#Tidying Up the Data
Now that we have our two data sets merged, we can now tidy our data so that it fits the 3 conventions of the tidyverse: (1) each variable has its own column, (2) each observation has its own row, (3) each value has its own cell.

We'll start off by giving each value its own cell and ensuring all observations are consistent by converting any plural observations to their singular forms. 
```{r}
phlox_clean_up1 <- phlox_merge %>%
  mutate_at("color", funs(str_replace(., " or", ", "))) %>% #mutations 1-4 are used to ensure all cells with more than one value are using the same separators, a comma
  mutate_at("color", funs(str_replace(., " and", ", "))) %>%
  mutate_at("pollinator", funs(str_replace(., " and", ", "))) %>%
  mutate_at("color", funs(str_replace(., "/", ", "))) %>%
  mutate_at("color", funs(str_replace(., " to", ", "))) %>%
  separate(color, into=c("color_1", "color_2", "color_3", "color_4"), sep=", ") %>% #there are a maximum of 4 different colors in one cell under the color column, so we will separate the original color column into 4 different rows to give each recorded color its own cell; NAs are put in the respective columns for species that have less than 4 different recorded colors. 
  mutate_all(funs(str_replace(., " primary,", ", "))) %>% #mutations 5-6 serve to remove extraneous verbiage that won't be relevant to our analysis
  mutate_all(funs(str_replace(., " secondary", ""))) %>%
  mutate_all(funs(str_replace(., "bees", "bee"))) %>% #mutations 7-26 serve to make our pollinator data consistent and turn any plural pollinators into singular, resolve inconsistencies in spacing, etc.
  mutate_all(funs(str_replace(., "butterflies", "butterfly"))) %>%
  mutate_all(funs(str_replace(., " butterfly", "butterfly"))) %>% #keep repeated mutations, otherwise it will count butterfly twice in n_distinct()
  mutate_all(funs(str_replace(., " butterfly", "butterfly"))) %>%
  mutate_all(funs(str_replace(., "beeflies", "bee-fly"))) %>%
  mutate_all(funs(str_replace(., "beefly", "bee-fly"))) %>%
  mutate_all(funs(str_replace(., "  bee-fly", "bee-fly"))) %>%
  mutate_all(funs(str_replace(., " bee-fly", "bee-fly"))) %>%
  mutate_all(funs(str_replace(., "beetles", "beetle"))) %>% 
  mutate_all(funs(str_replace(., " beetles", "beetle"))) %>%
  mutate_all(funs(str_replace(., "hawkmoths", "hawkmoth"))) %>% 
  mutate_all(funs(str_replace(., "  hawkmoth", "hawkmoth"))) %>% #keep repeated mutations, otherwise it will count hawkmoth twice in n_distinct()
  mutate_all(funs(str_replace(., " hawkmoth", "hawkmoth"))) %>%
  mutate_all(funs(str_replace(., "  bee", "bee"))) %>%
  mutate_all(funs(str_replace(., "flies", "fly"))) %>% 
  mutate_all(funs(str_replace(., " fly", "fly"))) %>% #keep repeated mutations, otherwise it will count fly twice in n_distinct()
  mutate_all(funs(str_replace(., " fly", "fly"))) %>%
  mutate_all(funs(str_replace(., "hummingbirds", "hummingbird"))) %>% 
  mutate_all(funs(str_replace(., " hummingbird", "hummingbird"))) %>% #keep repeated mutations, otherwise it will count hummingbird twice in n_distinct()
  mutate_all(funs(str_replace(., " hummingbird", "hummingbird"))) %>%
  separate(pollinator, into=c("pollinator_1", "pollinator_2", "pollinator_3"), sep=",") #there are a maximum of 3 different pollinators in one cell under the pollinator column, so we will separate the original pollinator column into 3 different rows to give each value its own cell; NAs are put in the respective columns for species that have less than 3 different recorded pollinators. 
phlox_clean_up1
```

We'll then do a more thorough clean-up under the color columns, making sure that all data is consistent here, as well. We looked to the paper for this study to make decisions on how to group/standardize recorded colors. We settled on the following categories which can then be correlated/attributed to pigments: blue_purple, pink, red (anthocyanins), yellow (carotenoids), green (chloroplasts), and white (lack of pigment). We also retained "bluish_white" and "greenish_yellow" as separate because they crossed over between two different pigment groups. 

```{r}
phlox_tidy <- phlox_clean_up1 %>%
  mutate_all(funs(str_replace(., "light pink", "pink"))) %>%
  mutate_all(funs(str_replace(., "pale pink", "pink"))) %>%
  mutate_all(funs(str_replace(., "pink,", "pink"))) %>%
  mutate_all(funs(str_replace(., "yellow throat", "yellow"))) %>%
  mutate_all(funs(str_replace(., "lavender", "blue_purple"))) %>%
  mutate_all(funs(str_replace(., "blue-violet", "blue_purple"))) %>%
  mutate_all(funs(str_replace(., "pale violet", "blue_purple"))) %>%
  mutate_all(funs(str_replace(., "violet", "blue_purple"))) %>%
  mutate_all(funs(str_replace(., "blue-purple", "blue_purple"))) %>%
  mutate_all(funs(str_replace(., "purple-blue", "blue_purple"))) %>%
  mutate_all(funs(str_replace(., "light purple", "blue_purple"))) %>%
  mutate_all(funs(str_replace(., "pale purple", "blue_purple"))) %>%
  mutate_all(funs(str_replace(., "pinkish purple", "blue_purple"))) %>%
  mutate_all(funs(str_replace(., "pale blue", "blue_purple"))) %>% 
  mutate_all(funs(str_replace(., "light blue", "blue_purple"))) %>%
  mutate_all(funs(str_replace(., "purple", "blue_purple"))) %>% 
  mutate_all(funs(str_replace(., "blue", "blue_purple"))) %>% 
  mutate_all(funs(str_replace(., "blue_purple_blue_purple", "blue_purple"))) %>%
  mutate_all(funs(str_replace(., "blue_purple_purple", "blue_purple"))) %>%
  mutate_all(funs(str_replace(., "bluish-white", "bluish_white"))) %>%
  mutate_all(funs(str_replace(., "greenish-yellow", "greenish_yellow"))) %>%
  mutate_all(funs(str_replace(., " blue_purple", "blue_purple"))) %>%
  mutate_all(funs(str_replace(., " pink", "pink"))) %>%
  mutate_all(funs(str_replace(., " white", "white"))) %>%
  mutate_all(funs(str_replace(., " [(]both[)]", "")))
phlox_tidy
```

The pivot_longer() command will condense our dataframe so that, while still giving each value its own cell, all colors will be under a single column and all pollinators will be under its own single column, as opposed to separated out into colors1-4 and, and pollinators 1-3. We removed the columns "color_number" and "pollinator_number" as they are unncecessary. The object `phlox_long` is what we will now use for our analyses.
```{r}
phlox_long <- phlox_tidy %>% 
  pivot_longer(color_1:color_4,
               names_to = "color_number",
               values_to = "color",
               values_drop_na = TRUE) %>% 
  pivot_longer(pollinator_1:pollinator_3,
               names_to = "pollinator_number",
               values_to = "pollinator",
               values_drop_na = TRUE) %>%
  select(-color_number, -pollinator_number)
phlox_long
```

```{r}
n_distinct(phlox_long$species) #Checking to see that we still have 145 distinct species that we're looking at. 
```

Adding genus column data:
```{r}
genus_name <- phlox_long$genus <- word(phlox_long$species, 1)
phlox_long %>% 
  tabyl(genus)
phlox_long %>% 
  count(genus)
#Dataframe1$COL2 <- word(Dataframe2$COL1, 1)
```
#The problem that I'm seeing here is that now R thinks that every repeated instance of a species in a genus counts as a distinct instance of that genus! How do we get it to reflect the appropriate counts/percentages of each genus? i.e. I want a count of each genus, under the condition that it's distinguishing between distinct species, and not just adding up all the repeats of each species.


#It's Plotty Time!

##Color Analysis
```{r}
phlox_long %>% 
  count(color)
```

##How do we get the colors and pollinators to be counted only once for distinct species? 
```{r}
phlox_long %>%
  filter(color!="NA") %>%
  ggplot(aes(x=color, fill=color)) +
  geom_bar(color = "black", show.legend = FALSE) + #removed legend because it just repeats what the x-axis says 
  scale_fill_manual(values = c("blue_purple" = "slateblue",
                               "bluish_white" = "cadetblue1",
                               "green" = "olivedrab",
                               "greenish_yellow" = "greenyellow",
                               "pink" = "pink",
                               "red" = "brown2",
                               "white" = "white",
                               "yellow" = "gold")) +
  labs(title = "Phlox Corolla Colors", x = "Corolla Color", y = "Number of Phlox Species")+
  scale_color_grey() + theme_classic()+
  theme(plot.title = element_text(size = rel(1.5), hjust = .5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title=element_text(size = rel(1.5), face="bold", hjust=.5))
```

##Pollinator Analysis
```{r}
phlox_long %>% 
  filter(pollinator!="") %>%
  count(pollinator)
```

```{r}
phlox_long %>% 
  filter(pollinator!="NA" & pollinator!="") %>%
  ggplot(aes(x=pollinator, fill=pollinator))+
  geom_bar(color = "black", show.legend = FALSE)+
  scale_color_grey() + theme_classic()+
  scale_fill_manual(values = c("hummingbird" = "slateblue",
                               "hawkmoth" = "azure2",
                               "fly" = "chartreuse2",
                               "butterfly" = "pink",
                               "beetle" = "greenyellow",
                               "bee-fly" = "maroon",
                               "bee" = "yellow",
                               "bat" = "midnightblue",
                               "autogamous" = "mediumorchid1")) +
  labs(title = "Pollinators of Phlox",
       x = "Pollinator",
       y = "Number of Phlox Species") +
  theme(plot.title = element_text(size = rel(1.5), hjust = .5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title=element_text(size = rel(1.5), face="bold", hjust=.5))
#Feel free to play with aesthetics
```

#Looking at Genera With Our Continuous Variables of Mean Corolla Length, Mean Corolla Throat Width, and Mean Corolla Length:Width Ratio Using Box Plots:

##Mean Corolla Length (cm)
```{r}
min(phlox_long$mean_corolla_length_cm)
max(phlox_long$mean_corolla_length_cm)
class(phlox_long$mean_corolla_length_cm)
phlox_long$mean_corolla_length_cm <- as.numeric(phlox_long$mean_corolla_length_cm)
is.numeric(phlox_long$mean_corolla_length_cm)
```

```{r}
phlox_long %>% 
  ggplot(aes(x=genus, y=mean_corolla_length_cm))+
  geom_boxplot()+
  scale_color_grey() + theme_classic()+
  labs(title = "Corolla Lengths by Genus",
       x = "Phlox Genus",
       y = "Corolla Length (cm)") +
  theme(plot.title = element_text(size = rel(1.5), hjust = .5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title=element_text(size = rel(1.5), face="bold", hjust=.5))
```
##Mean Corolla Throat Width (cm)
```{r}
class(phlox_long$mean_corolla_width_throat_cm)
phlox_long$mean_corolla_width_throat_cm <- as.numeric(phlox_long$mean_corolla_width_throat_cm)
class(phlox_long$mean_corolla_width_throat_cm)
min(phlox_long$mean_corolla_width_throat_cm)
max(phlox_long$mean_corolla_width_throat_cm)
```
##Mean Corolla Throat Width (cm)
```{r}
phlox_long %>% 
  ggplot(aes(x=genus, y=mean_corolla_width_throat_cm))+
  geom_boxplot()+
  scale_y_log10() +
  scale_color_grey() + theme_classic()+
  labs(title = "Corolla Throat Widths by Genus",
       x = "Phlox Genus",
       y = "Corolla Width (log10 cm)") +
  theme(plot.title = element_text(size = rel(1.5), hjust = .5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title=element_text(size = rel(1.5), face="bold", hjust=.5))
```
##Corolla Length:Width Ratio 
```{r}
class(phlox_long$mean_length_width_ratio)
phlox_long$mean_length_width_ratio <- as.numeric(phlox_long$mean_length_width_ratio)
class(phlox_long$mean_length_width_ratio)
min(phlox_long$mean_length_width_ratio)
max(phlox_long$mean_length_width_ratio)
```

```{r}
phlox_long %>% 
  ggplot(aes(x=genus, y=mean_length_width_ratio))+
  geom_boxplot()+
  scale_color_grey() + theme_classic()+
  labs(title = "Corolla Length to Width Ratios by Genus",
       x = "Phlox Genus",
       y = "Corolla Length:Width") +
  theme(plot.title = element_text(size = rel(1.5), hjust = .5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title=element_text(size = rel(1.5), face="bold", hjust=.5))
```
```{r}
phlox_long %>% 
  filter(pollinator != "") %>% 
  ggplot(aes(x=mean_corolla_length_cm, y=mean_corolla_width_throat_cm))+
  geom_jitter()+
  geom_smooth(method = "lm")+
  theme_classic()+
  labs(title = "Relationship between Corolla Length and Corolla Width",
       x = "Mean Corolla Length (cm)",
       y = "Mean Corolla Width (cm)") +
  theme(plot.title = element_text(size = rel(1.5), hjust = .5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title=element_text(size = rel(1.5), face="bold", hjust=.5))
```

#Let's do plots comparing color and pollinators to corolla length:width ratio.

##Color and Corolla Length:Width Ratio
```{r}
phlox_long %>% 
  ggplot(aes(x=color, y=mean_length_width_ratio))+
  geom_boxplot()+
  scale_color_grey() + theme_classic()+
  labs(title = "Corolla Length to Width Ratios by Color",
       x = "Corolla Color",
       y = "Corolla Length:Width") +
  theme(plot.title = element_text(size = rel(1.5), hjust = .5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title=element_text(size = rel(1.5), face="bold", hjust=.5))
```
##Pollinator and Corolla Length:Width Ratio
```{r}
phlox_long %>% 
  filter(pollinator != "") %>% 
  ggplot(aes(x=pollinator, y=mean_length_width_ratio))+
  geom_boxplot()+
  scale_color_grey() + theme_classic()+
  labs(title = "Corolla Length to Width Ratios by Pollinator",
       x = "Pollinator",
       y = "Corolla Length:Width") +
  theme(plot.title = element_text(size = rel(1.5), hjust = .5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title=element_text(size = rel(1.5), face="bold", hjust=.5))
```

#Relative Frequency of Corolla Colors Among Pollinators:  
```{r}
phlox_long %>% 
  filter(pollinator!="") %>%
  #filter(color!="NA") %>%
  ggplot(aes(x = pollinator, fill = color)) + geom_bar(position = position_fill(), color = "black") +
  scale_color_grey() + theme_classic()+
  scale_fill_manual(values = c("blue_purple" = "slateblue",
                               "bluish_white" = "cadetblue1",
                               "green" = "olivedrab",
                               "greenish_yellow" = "greenyellow",
                               "pink" = "pink",
                               "red" = "brown2",
                               "white" = "white",
                               "yellow" = "gold")) +
  labs(title = "Pollinators' Preferred Flower Color",
       x = "Pollinator",
       fill = "color")+
   theme(plot.title = element_text(size = rel(1.5), hjust = .5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title=element_text(size = rel(1.5), face="bold", hjust=.5))
```
We can start to see some trends in phlox data - we can see that phlox pollinators like hummingbird are overwhelmingly attracted to red flowers. Bees and flies are mostly attracted to the blue_purple range. Bats are mostly attracted to green, but will also pollinate pink or blue_purple.

#Looking at which genera are pollinated by which pollinators: 
```{r}
phlox_long %>% 
  count(genus)
```
```{r}
phlox_long %>% 
  filter(pollinator!="") %>%
  ggplot(aes(x = genus, fill = pollinator)) + geom_bar(position = position_fill(), color = "black") +
  scale_color_grey() + theme_classic()+
  scale_fill_manual(values = c("hummingbird" = "slateblue",
                               "hawkmoth" = "azure2",
                               "fly" = "chartreuse2",
                               "butterfly" = "pink",
                               "beetle" = "greenyellow",
                               "bee-fly" = "maroon",
                               "bee" = "yellow",
                               "bat" = "midnightblue",
                               "autogamous" = "mediumorchid1")) +
  labs(title = "Pollinators of Phlox Genera",
       x = "Phlox Genus",
       fill = "color")+
   theme(plot.title = element_text(size = rel(1.5), hjust = .5))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title=element_text(size = rel(1.5), face="bold", hjust=.5))
```
#It should be noted, that some genera are only represented by one species so we can't draw any definite conclusions about some genera only being pollinated by one type of pollinator. 

App Looking at Continuous Variables in the Plot:
```{r}
ui <- fluidPage(
    selectInput("x", "Select X Variable", choices = c("mean_corolla_length_cm", "mean_corolla_width_throat_cm", "mean_length_width_ratio"),
              selected = "mean_corolla_length_cm"),
    selectInput("y", "Select Y Variable", choices = c("mean_corolla_length_cm", "mean_corolla_width_throat_cm", "mean_length_width_ratio"),
              selected = "mean_corolla_width_throat_cm"),
  plotOutput("plot", width = "1200px", height = "400px")
)

server <- function(input, output) {
  output$plot <- renderPlot({
    ggplot(phlox_long, aes_string(x = input$x, y = input$y, color="genus")) + 
      geom_point() + 
      scale_color_grey() + theme_classic()+
      theme_light(base_size = 18)+
      scale_x_discrete(breaks = seq(0, 10, 1))+
      scale_y_discrete(breaks = seq(0, 10, 1))
  })
  session$onSessionEnded(stopApp)
} #need to play with aesthetics so that legend has more distinct colors

shinyApp(ui, server)
```

choices = c("Acanthogilia", "Aliciella", "Allophyllum", "Bonplandia", "Cantua", "Cobaea", "Collomia", "Dayia", "Eriastrum", "Fouquieria", "Gilia", "Giliastrum", "Gymnosteris", "Ipomopsis", "Langloisia", "Leptosiphon", "Linanthus", "Loeselia", "Loeseliastrum", "Microgilia", "Microsteris", "Navarretia", "Phlox", "Polemonium", "Saltugilia")


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
